Accelerated oblique random survival forests

### 2022-05-24 paper/jmlr/main.Rnw
much more modular sim

\usepackage{jmlr2e}
% extra packages for tables
\usepackage{longtable}
\usepackage{colortbl}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{caption}

@@

Several machine learning algorithms can engage with right-censored time-to-event outcomes. In the current study, we consider four classes of learners: random forests, boosting ensembles, regression models, and neural networks.

@@

@@

results_overall <- bm_pred_comb |>
  group_by(model) |>
  summarize(
    across(
      .cols = c(cstat, IPA, time_fit, time_prd),
      .fns = list(mean = mean, sd = sd)
    )
  ) |>
  mutate(data = 'Overall')

data_tbl <- bm_pred_comb |>
  group_by(model, data, n_z, n_obs, correlated_x) |>
  summarize(
    across(
      .cols = c(cstat, IPA, time_fit, time_prd),
      .fns = list(mean = mean, sd = sd)
    )
  ) |>
  bind_rows(results_overall) |>
  mutate(
    data = if_else(
      is.na(n_z),
      true = data,
      false = as.character(
        glue::glue("Simulation, N junk = {n_z}, N obs = {n_obs}, X corr less then or equal to {correlated_x}")
      )
    )
  ) |>
  group_by(data) |>
  mutate(
    across(.cols = starts_with("time"), .fns = as.numeric),
    time_fit_ratio = time_fit_mean / time_fit_mean[model == 'aorsf-1'],
    time_prd_ratio = time_prd_mean / time_prd_mean[model == 'aorsf-1']
  ) |>
  arrange(data, desc(cstat_mean)) |>
  transmute(
    model = recode(
      model,
      "aorsf-15" = 'aORSF(i=15)',
      "aorsf-1" = 'aORSF(i=1)',
      "ranger" = 'Ranger',
      "xgboost" = 'Xgboost',
      "cif" = "Party",
      "randomForestSRC" = "Rfsrc",
      "coxtime" = "Coxtime"
    ),
    data = recode(
      data,
      'pbc_orsf' = 'Mayo Clinic Primary Biliary Cholangitis Data, N = 276',
      'rotterdam' = 'Rotterdam Breast Cancer Data, N = 2,982',
      'sim' = 'Simulated data'
    ),
    cstat = table_glue("{cstat_mean} ({cstat_sd})"),
    ipa = table_glue("{IPA_mean} ({IPA_sd})"),
    time_fit_mean = format(round(time_fit_mean, 3), nsmall = 3),
    time_fit_ratio = table_value(time_fit_ratio),
    time_prd_mean = format(round(time_prd_mean, 3), nsmall = 3),
    time_prd_ratio = table_value(time_prd_ratio)
  )

>>>>

\usepackage{jmlr2e}
% extra packages for tables
\usepackage{longtable}
\usepackage{colortbl}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{hyperref}
\hypersetup{hidelinks}

@@

Table \ref{table:1} is an example of a referenced \LaTeX{}. Several machine learning algorithms can engage with right-censored time-to-event outcomes. In the current study, we consider four classes of learners: random forests, boosting ensembles, regression models, and neural networks.

@@

\begin{table}[h!]
\centering
\begin{tabular}{p{2cm} | p{3cm} p{7cm}}
 \hline
 Learner & Software & Description \\ [0.5ex]
 \hline\hline
 \multicolumn{3}{l}{\textit{Random Survival Forests}}\\
 \hline\hline
 Standard & \href{https://www.randomforestsrc.org/index.html}{\texttt{RandomForestSRC}} & Axis based survival trees following Leo Breiman's original random forest algorithm, with cut-points selected to maximize a log-rank statistic.  \\ \hline
 Oblique & \href{https://CRAN.R-project.org/package=obliqueRSF}{\texttt{obliqueRSF}} \newline \href{https://bcjaeger.github.io/aorsf/}{\texttt{aorsf}} & Oblique survival trees following Leo Breiman's random forest algorithm but not using random coefficients for linear combinations. \texttt{obliqueRSF} uses penalized models, and \texttt{aorsf} additionally uses partially trained models. Cut-points are selected to maximize a log-rank statistic. \\ \hline
 Extremely \,Randomized & \href{https://CRAN.R-project.org/package=ranger}{\texttt{ranger}} & Axis-based survival trees grown with randomly selected features and cut-points\\ \hline
 Conditional \,Inference & \href{http://party.r-forge.r-project.org/}{\texttt{party}} & Axis based survival trees grown using unbiased recursive partitioning.  \\
 \hline\hline
 \multicolumn{3}{l}{\textit{Boosting ensembles}}\\
 \hline\hline
 Trees & \href{https://xgboost.readthedocs.io/en/stable/#}{\texttt{xgboost}} & The Cox likelihood function is maximized additively with decision trees. Nested cross validation (5 folds) is applied to tune the number of trees grown.  \\
 Models & \href{https://xgboost.readthedocs.io/en/stable/#}{\texttt{xgboost}} & The accelerated failure time likelihood function is maximized additively with decision trees. Nested cross validation (5 folds) is applied to tune the number of trees grown.  \\
 \hline\hline
 \multicolumn{3}{l}{\textit{Regression models}}\\
 \hline\hline
 Cox Net & \texttt{glmnet} & The Cox model is fit using an elastic net penalty. Nested cross validation (5 folds) is applied to tune penalty terms.\\
 \hline\hline
 \multicolumn{3}{l}{\textit{Neural networks}}\\
 \hline\hline
 Cox Time & 6 & 87837  \\
 \hline
\end{tabular}
\caption{Table to test captions and labels.}
\label{table:1}
\end{table}

@@

results_overall <- bm_pred_comb |>
  group_by(model) |>
  summarize(
    across(
      .cols = c(cstat, IPA, time_fit, time_pred),
      .fns = list(mean = mean, sd = sd)
    )
  ) |>
  mutate(data = 'Overall')

data_tbl <- bm_pred_comb |>
  group_by(model, data, n_z, n_obs, correlated_x) |>
  summarize(
    across(
      .cols = c(cstat, IPA, time_fit, time_pred),
      .fns = list(mean = mean, sd = sd)
    )
  ) |>
  bind_rows(results_overall) |>
  mutate(
    data = if_else(
      is.na(n_z),
      true = data,
      false = as.character(
        glue::glue("Simulation, N junk = {n_z}, N obs = {n_obs}, X corr less then or equal to {correlated_x}")
      )
    )
  ) |>
  group_by(data) |>
  mutate(
    across(.cols = starts_with("time"), .fns = as.numeric),
    time_fit_ratio = time_fit_mean / time_fit_mean[model == 'aorsf_1'],
    time_pred_ratio = time_pred_mean / time_pred_mean[model == 'aorsf_1']
  ) |>
  arrange(data, desc(cstat_mean)) |>
  transmute(
    model = recode(
      model,
      "aorsf_15" = 'aORSF(i=15)',
      "aorsf_1" = 'aORSF(i=1)',
      "ranger" = 'Ranger',
      "xgboost" = 'Xgboost',
      "cif" = "Party",
      "randomForestSRC" = "Rfsrc",
      "coxtime" = "Coxtime"
    ),
    data = recode(
      data,
      'pbc_orsf' = 'Mayo Clinic Primary Biliary Cholangitis Data, N = 276',
      'rotterdam' = 'Rotterdam Breast Cancer Data, N = 2,982',
      'sim' = 'Simulated data'
    ),
    cstat = table_glue("{cstat_mean} ({cstat_sd})"),
    ipa = table_glue("{IPA_mean} ({IPA_sd})"),
    time_fit_mean = format(round(time_fit_mean, 3), nsmall = 3),
    time_fit_ratio = table_value(time_fit_ratio),
    time_pred_mean = format(round(time_pred_mean, 3), nsmall = 3),
    time_pred_ratio = table_value(time_pred_ratio)
  )
